---
title: "Historical CEDS vs default Hector"
output: html_document
date: "`r Sys.Date()`"
---

```{r knit settings, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      out.width='80%', fig.align='center', class='center', 
                      fig.path = "figs/")

BASE_DIR <- here::here()

library(dplyr)
library(hector)
library(ggplot2)
theme_set(theme_bw())
```

Here we would like to do a comparison of the emission inputs we have created with CEDS and the default total emissions. This will help us identify where there are potential sectors missing and/or issues with emission conversion factors. We include a vertical line at at the 2015, the final year of the open burning emissions data set which was extended until 2022 but holding the 2015 values constant until 2022. 


Some lingering questions 

(1) Around 2000 the emissions for BC, NOx, OC emissions diverge, how did the CEDS emissions change over those time periods? Between the different versions? 
(2) For the non CEDS emissions (LUC emissions, halocarbon, ect.) would we like to use the SSP2-45 emissions? Or is there a different scenario we should consider? 
(3) How woudl we like to transition from the historical to the future period? 


```{r}
# Import the new historical emissions based on the new CEDS release
read.csv(file.path(BASE_DIR, "data", "L2", "hector_ceds.csv")) %>% 
  mutate(source = "new") -> 
  new_emissions

# Load the old historical emissions that are used as comparison 
read.csv(file.path(BASE_DIR, "data", "L0", "L0.V32_emissions.csv")) %>% 
  mutate(source = "old") -> 
  old_emissions

new_emissions %>% 
  bind_rows(old_emissions) %>% 
  filter(variable %in% new_emissions$variable) %>% 
  select(year, variable, units, value, source) -> 
  emiss_df
```


```{r}

# Define a helper function to plot the new vs old emission time series against one another
# Args 
#   emiss_df: data.frame of the emissions, it can contain results for multiple emission types 
#   var: string, variable name to plot 
# Returns a ggplot figure
quick_emiss_plot <- function(emiss_df, var){
  
  unit_label <- getunits(var)

  emiss_df %>% 
    filter(variable == var) %>% 
    ggplot() + 
    geom_vline(xintercept = 2016, color = "grey")+ 
    geom_line(aes(year, value, color = source), linewidth = 0.75) + 
    labs(title = var, y = unit_label, x = NULL) + 
    scale_color_manual(values = c("old" = "darkgrey", "new" = "blue")) -> 
    plot
  
  return(plot)
}

```


A of `r Sys.Date()` this includes only CEDS + open burning emissions

 
# BC Emissions

```{r bc_plot}
quick_emiss_plot(var = EMISSIONS_BC(), emiss_df)
```

# CH4 Emissions

```{r ch4_plot}
quick_emiss_plot(var = EMISSIONS_CH4(), emiss_df)
```


# CO Emissions

```{r, co_plot}
quick_emiss_plot(var = EMISSIONS_CO(), emiss_df)
```


# N2O Emissions

```{r n2o_plot}
quick_emiss_plot(var = EMISSIONS_N2O(), emiss_df)
```

# NH3 Emissions
```{r nh3_plot}
quick_emiss_plot(var = EMISSIONS_NH3(), emiss_df)
```

# NMVOC Emissions
```{r nmvoc_plot}
quick_emiss_plot(var = EMISSIONS_NMVOC(), emiss_df)
```

# NOx Emissions
```{r nox_plot}
quick_emiss_plot(var = EMISSIONS_NOX(), emiss_df)
```

# OC Emissions 
```{r oc_plot}
quick_emiss_plot(var = EMISSIONS_OC(), emiss_df)
```

# SO2 Emissions
```{r so2_plot}
quick_emiss_plot(var = EMISSIONS_SO2(), emiss_df)
```

# FFI Emissions
```{r ffi_plot}
quick_emiss_plot(var = FFI_EMISSIONS(), emiss_df)
```