---
title: "Benchmarking RCMIP L1 Outputs"
author: "Kalyn Dorheim"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true             # Enables the Table of Contents
    toc_float:            # Configures the floating TOC
      collapsed: false    # Whether the TOC starts collapsed
      smooth_scroll: true # Enables smooth scrolling to sections
---

This is some extra QA/QC to to make sure that we are not introducing new errors by 
do the RCMIP conversions incorrectly. 

## Set Up 
```{r, message=FALSE, warning=FALSE}
# Load required libraries 
library(dplyr)
library(ggplot2)
library(hector)
library(tidyr)

# Set the graphics theme and define the base relative pathway 
theme_set(theme_bw())
BASEDIR <- here::here()
```

```{r}
# Read in the "new" rcmip emissions aka the emissions that are being calculated 
# with the workflow. 
read.csv(file.path(BASEDIR, "data", "L1", "L1.hector_rcmip_emiss.csv")) %>% 
    mutate(source = "new") %>% 
    select(-units) -> 
    new_emiss
```


```{r}
# Save all the scenario names, these strings will be used to load the default emissions 
# from the Hector package. 
scns <- setdiff(unique(new_emiss$scenario), "historical")

system.file(package = "hector", "input/tables") %>% 
    list.files(pattern = paste0(scns, collapse = "|"), full.names = TRUE) %>% 
    lapply(function(f){
        
        scn <- gsub(pattern = "_emiss-constraints_rf.csv", replacement = "", x = basename(f))
        d <- read.csv(f, comment.char = ";")
        
        d %>% 
            pivot_longer(cols = -Date, names_to = "variable", values_to = "value") %>%  
            mutate(scenario = scn, 
                   source = "default") -> 
            long_df 
        
        return(long_df)
        
    }) %>% 
    do.call(what = "rbind") %>%  
    rename(year = Date) -> 
    default_emiss
```

```{r}
emiss_long <- rbind(default_emiss, new_emiss)
```

# Summary Stats on Differences 

```{r}
# Define our acceptable tolerance of differences aka what do we consider to be 0? 
tol <- 1e-4

```


```{r}
default_emiss %>% 
    select(scenario, year, default = value, variable) %>% 
    inner_join(
        new_emiss %>% 
            select(scenario, year, new = value, variable), by = c("scenario", "year", "variable")) %>% 
    mutate(abs_dif = abs(new - default), 
           rel_dif = abs(new - default)/abs(default)) -> 
    diff_table
```


```{r}
diff_table %>% 
    group_by(scenario, variable) %>% 
    summarise(total = sum(abs_dif)) -> 
    total_dif

# So it does look like there are some major difference between some of the emission species... 
diff_table %>%  
    filter(abs_dif > tol) %>% 
    pull(variable) %>% 
    unique()
```

Initially there were some emissions that had some issues (differences between the new and the default) so we took a deeper look at them in the following sections. 


# CH$_4$ 

How different are the CH$_4$ results? - Not much they are essentially identical 

```{r}
emiss_long %>% 
    filter(variable == EMISSIONS_CH4()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```


Answer - Not much any more 

# N$_2$O 


How different are the N$_2$O results? - Not much they are essentially identical 

```{r}
total_dif %>% 
    filter(variable == EMISSIONS_N2O())
```

```{r}
diff_table %>% 
    filter(variable == EMISSIONS_N2O()) %>% 
    filter(abs_dif > tol) %>% 
    group_by(scenario) %>% 
    summarise(min = min(abs_dif), 
              mean = mean(abs_dif), 
              max = max(abs_dif))
```




```{r}
emiss_long %>% 
    filter(variable == EMISSIONS_N2O()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```



# NMVOC 

How different are the NMVOC results? 


```{r}
emiss_long %>% 
    filter(variable == EMISSIONS_NMVOC()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```
# SO$_2$ 

How different are the SO$_2$  results? - Not different, for our purposes identical


```{r}
diff_table %>% 
    filter(variable == EMISSIONS_SO2()) %>% 
    filter(abs_dif > tol) 
```

```{r}
emiss_long %>% 
    filter(variable == EMISSIONS_SO2()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```


# NO$_x$ 

How different are the NO$_x$  results? - Not different, for our purposes identical


```{r}
diff_table %>% 
    filter(variable == EMISSIONS_NOX()) %>% 
   # filter(abs_dif > tol) %>% 
    group_by(scenario) %>% 
    summarise(min = min(abs_dif), 
              mean = mean(abs_dif), 
              max = max(abs_dif))
```

```{r}
emiss_long %>% 
    filter(variable == EMISSIONS_NOX()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```
# LUC emissions 

How different are the luc emissions? - they are pretty much the same 


```{r}
diff_table %>% 
    filter(variable == LUC_EMISSIONS()) %>% 
   # filter(abs_dif > tol) %>% 
    group_by(scenario) %>% 
    summarise(min = min(abs_dif), 
              mean = mean(abs_dif), 
              max = max(abs_dif))
```

```{r}
emiss_long %>% 
    filter(variable == LUC_EMISSIONS()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```


# LUC uptake 

How different are the luc uptake? - they are pretty much the same 


```{r}
diff_table %>% 
    filter(variable == LUC_UPTAKE()) %>% 
   filter(abs_dif > tol) %>% 
    group_by(scenario) %>% 
    summarise(min = min(abs_dif), 
              mean = mean(abs_dif), 
              max = max(abs_dif))
```

```{r}
emiss_long %>% 
    filter(variable == LUC_UPTAKE()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```
# FFI Emissions 

How different are the ffi emissions? - they are pretty much the same 


```{r}
diff_table %>% 
    filter(variable == FFI_EMISSIONS()) %>% 
   filter(abs_dif > tol) %>% 
    group_by(scenario) %>% 
    summarise(min = min(abs_dif), 
              mean = mean(abs_dif), 
              max = max(abs_dif))
```

```{r}
emiss_long %>% 
    filter(variable == FFI_EMISSIONS()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```

# DACCS Uptake 

How different are the daccs uptake? - they are pretty much the same 


```{r}
diff_table %>% 
    filter(variable == DACCS_UPTAKE()) %>% 
   filter(abs_dif > tol) %>% 
    group_by(scenario) %>% 
    summarise(min = min(abs_dif), 
              mean = mean(abs_dif), 
              max = max(abs_dif))
```

```{r}
emiss_long %>% 
    filter(variable == DACCS_UPTAKE()) %>% 
    ggplot(aes(year, value, color = source, linetype = source)) + 
    geom_line() +
    facet_wrap("scenario")
```
