---
title: "Benchmarking L1 Outputs"
author: "Kalyn Dorheim"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true             # Enables the Table of Contents
    toc_float:            # Configures the floating TOC
      collapsed: false    # Whether the TOC starts collapsed
      smooth_scroll: true # Enables smooth scrolling to sections
---

Here I compare the L1 outputs (CEDs + the burning emissions from RCMIP) with the default emissions passed into Hector. 

```{r, message=FALSE, warning=FALSE}
# Packages used here 
library(dplyr)
library(ggplot2)
library(hector)
library(kableExtra)
library(knitr)
library(tidyr)

# Using relative paths! 
BASE_DIR <- here::here()

# Set the default plot optics
theme_set(theme_bw())
```


```{r}
# Import CEDs data! 
file.path(BASE_DIR, "data", "L1", "L1.incomplete_ceds_hector.csv") %>%  
    read.csv() %>% 
    mutate(source = "CEDs +") -> 
    ceds_emiss

# Run Hector emissions to compare with CEDS 
scns <- c("ssp119", "ssp245", "ssp585")
vars <- unique(ceds_emiss$variable)
yrs <- min(ceds_emiss$year):(max(ceds_emiss$year) + 20)

lapply(scns, function(scn){
    ini_file <- list.files(system.file(package = "hector", "input"), 
                           pattern = scn, full.names = TRUE)
    hc <- newcore(ini_file, name = scn)
    run(hc)
    fetchvars(core = hc, dates = yrs, vars = vars) %>%  
        mutate(source = "hector default") ->
        out

    return(out)
}) %>% 
    do.call(what = "rbind") -> 
    hector_emiss

hector_emiss %>% 
    filter(scenario == "ssp245") %>% 
    filter(year <= 2014) %>% 
    mutate(scenario = "historical") -> 
    hector_historical

all_emiss <- bind_rows(ceds_emiss, hector_emiss, hector_historical)
```

# Relative Differences
```{r}
all_emiss %>%  
    filter(scenario == "historical") %>% 
    mutate(source = if_else(source == "hector default", "H", "C")) %>% 
    select(year, value, variable, source) %>% 
    spread(source, value) %>% 
    na.omit %>% 
    mutate(diff = (C-H)/H) -> 
    differences
```


What is the relative difference between the default Hector and CEDS emissions? This will help us get a sense of what emissions have been converted properly and that the CEDS & RCMIP emissions are aggregated properly.

```{r}
differences %>% 
    group_by(variable) %>% 
    summarise(min = min(diff), 
              mean = mean(diff), 
              max = max(diff), 
              sd = sd(diff)) %>% 
  kable(digits = 4, format = "html", caption = "Summary Stats for Relative Difference by Emission") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "center"
  )
```

```{r}
ggplot(differences) + 
    geom_boxplot(aes(variable, diff)) + 
    labs(y = "(CEDS - Hector) / Hector", x = NULL) + 
      theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

# BC Emissions

CEDS+ = global total BC CEDS emissions +
    (Emissions|BC|MAGICC AFOLU|Agricultural Waste Burning,
    Emissions|BC|MAGICC AFOLU|Forest Burning, 
    Emissions|BC|MAGICC AFOLU|Grassland Burning,
    Emissions|BC|MAGICC AFOLU|Peat Burning)

```{r}
var <- EMISSIONS_BC()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2030) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario") + 
    labs(title = var, y = getunits(var))

```


# CH4 Emissions

CEDs+ = only CEDS, there were not additional CH4 burning emissions 

```{r}
var <- EMISSIONS_CH4()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year %in% 1950:2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```

# CO Emissions

CEDs+  = CO global total CO emissions +
        (Emissions|CO|MAGICC AFOLU|Forest Burning, 
        Emissions|CO|MAGICC AFOLU|Grassland Burning, 
        Emissions|CO|MAGICC AFOLU|Peat Burning, 
        Emissions|CO|MAGICC AFOLU|Agricultural Waste Burning)

```{r}
var <- EMISSIONS_CO()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```


# CO2 Emissions (ffi only)

CEDs+ is actually only CEDs

```{r}
var <- FFI_EMISSIONS()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```

# N2O Emissions 

CEDs + is only CEDs no emissions from RCMIP 

```{r}
var <- EMISSIONS_N2O()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year %in% 1900:2030) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))
```

# NH3 Emissions 

CEDS+ = global total NH3 CEDs emiss + 
            (Emissions|NH3|MAGICC AFOLU|Forest Burning
                Emissions|NH3|MAGICC, AFOLU|Grassland Burning,
                Emissions|NH3|MAGICC AFOLU|Peat Burning,
                Emissions|NH3|MAGICC AFOLU|Agricultural Waste Burning)
```{r}
var <- EMISSIONS_NH3()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```

# NMVOC Emissions 


CEDS+ = global total NMVOC CEDs emiss + 
            (Emissions|VOC|MAGICC AFOLU|Forest Burning
                Emissions|VOC|MAGICC, AFOLU|Grassland Burning,
                Emissions|VOC|MAGICC AFOLU|Peat Burning,
                Emissions|VOC|MAGICC AFOLU|Agricultural Waste Burning)


```{r}
var <- EMISSIONS_NMVOC()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))
```


# OC Emissions

CEDS+ = global total OC CEDs emiss + 
        (Emissions|VOC|MAGICC AFOLU|Forest Burning
            Emissions|OC|MAGICC, AFOLU|Grassland Burning,
            Emissions|OC|MAGICC AFOLU|Peat Burning,
            Emissions|OC|MAGICC AFOLU|Agricultural Waste Burning)

```{r}
var <- EMISSIONS_OC()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```


# NOx Emissions

CEDS+ = global total NOx CEDs emiss + 
    (Emissions|NOx|MAGICC AFOLU|Forest Burning
        Emissions|NOx|MAGICC AFOLU|Grassland Burning
        Emissions|NOx|MAGICC AFOLU|Peat Burning
        Emissions|NOx|MAGICC AFOLU|Agricultural Waste Burning)
```{r}
var <- EMISSIONS_NOX()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```


# S2O Emissions

CEDs+ = CEDs total SO2 emissions + 
            (Emissions|Sulfur|MAGICC AFOLU|Forest Burning + 
                Emissions|Sulfur|MAGICC AFOLU|Grassland Burning + 
                Emissions|Sulfur|MAGICC AFOLU|Peat Burning + 
                Emissions|Sulfur|MAGICC AFOLU|Agricultural Waste Burning)

```{r}
var <- EMISSIONS_SO2()

all_emiss %>% 
    filter(variable == var) %>% 
    filter(year <= 2022) %>% 
    ggplot(aes(year, value, color = source)) + 
    geom_line() +
    facet_wrap("scenario", scales = "free") + 
    labs(title = var, y = getunits(var))

```




