---
title: "Open Burning Comparison"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
```


# CMIP6 Era Open Burning Emissions

```{r}
# Load the BB4CMIP data files
# Args
#   f: string path to the data file to read in
# Returns: data.frame of global emission per sector
load_BB4CMIP_fxn <- function(f){
  
  # Each of the files corresponded to a type of emissions single sector.
  d <- read.csv(f)
  
  # assert that the data only contains one type of data
  ver <- unique(d$ver)
  stopifnot(length(ver) == 1)
  
  sector <- unique(d$sector)
  stopifnot(length(sector) == 1)
  
  # Determine emission species
  em <- unlist(strsplit(basename(f), split = "_|-"))[[1]]
  
  # Extract the global values
  inds <- which(grepl(x = colnames(d), pattern = "X"))
  value <- as.vector(colSums(d[, inds]))
  years <- as.vector(as.numeric(gsub(x = colnames(d)[inds], pattern = "X", replacement = "")))
  units <- "Tg"
  
  # Format output
  out <- data.frame(year = years, sector, variable = em, value, units)
  return(out)
  
  
}


here::here("scripts", "dev", "1750-2015_v1.2_Emissions_bulk_em") %>% 
  list.files(pattern = "csv", full.names = TRUE) -> 
  files

# Assert that all the files exists, if this throws an error 
# see the read me in the data-raw directory. 
stopifnot(all(file.exists(files)))

files %>% 
  lapply(load_BB4CMIP_fxn) %>%
  do.call(what = "rbind") %>% 
  summarise(value = sum(value), .by = c("year", "variable")) ->
  cmip6_emissions

```


# CMIP7 Era Open Burning Emissions

```{r}
# Load the BB4CMIP data files
# Args
#   f: string path to the data file to read in
# Returns: data.frame of global emission per sector
load_BB4CMIP7_fxn <- function(f){
  
  # Each of the files corresponded to a type of emissions single sector.
  d <- read.csv(f)
  
  # Make sure we have the global values 
  stopifnot(unique(d$region) == "World")
  
  
  
  # Determine emission species
  em <- unlist(strsplit(basename(f), split = "_|-"))[[1]]
  
  # Extract the global values
  inds <- which(grepl(x = colnames(d), pattern = "X"))
  
  # Format the output 
  pivot_longer(d, names_to = "year", values_to = "value", cols = inds) %>% 
    mutate(year = as.numeric(gsub(replacement = "", pattern = "X", x = year))) -> 
    out
  
  return(out)
  
  
}




# Import raw data files
here::here("scripts", "dev") %>% 
  list.files(pattern = "gfed-bb4cmip_cmip7_global_0011", full.names = TRUE) -> 
  files

# Assert that all the files exists, if this throws an error 
# see the read me in the data-raw directory. 
stopifnot(all(file.exists(files)))

files %>% 
  lapply(load_BB4CMIP7_fxn) %>%
  do.call(what = "rbind") %>% 
  mutate(variable = gsub(replacement = "", x = variable, pattern = "Emissions|Biomass Burning")) %>% 
  mutate(variable = gsub(replacement = "", x = variable, pattern = "\\|")) %>% 
  mutate(variable = gsub(replacement = "NMVOC", x = variable, pattern = "VOC")) %>% 
  mutate(variable = gsub(replacement = "SO2", x = variable, pattern = "Sulfur")) -> 
  cmip7_emissions

```


# Compare Open Burning 

Okay so right off the bat these are in different units... 

```{r}
cmip6_emissions$source <- "cmip6"
cmip7_emissions$source <- "cmip7"
raw_emissions <- bind_rows(cmip6_emissions, cmip7_emissions)

raw_emissions %>% 
  filter(variable == em) %>% 
  ggplot(aes(year, value, color = source)) + 
  geom_line() + 
  labs(title = "CMIP6 vs. CMIP7 Open Burning Emissions")
```


```{r}

cmip6_emissions %>% 
  mutate(value = value * 0.001, 
         source = "cmip6") -> 
  cmip6_emissions
cmip7_emissions$source <- "cmip7"
emissions <- bind_rows(cmip6_emissions, cmip7_emissions)


quick_plot <- function(emissions, em){
  
  emissions %>% 
    dplyr::filter(variable == em) -> 
    to_plot 
  
  to_plot %>%  
    summarise(min = min(value), 
              max = max(value), 
              mean = mean(value), 
              sd = sd(value), 
              .by = c("variable", "source")) -> 
    sum_stats 
  
  to_plot %>% 
    ggplot(aes(year, value, color = source), alpha = 0.5) + 
    geom_line() + 
    labs(title = "CMIP6 vs. CMIP7 Open Burning Emissions") -> 
    plot 
  
  out <- list(plot = plot, stats = sum_stats) 
  return(out)
  
}
```


```{r}
quick_plot(emissions, "CH4")
```

```{r}
quick_plot(emissions, "BC")
```

```{r}
quick_plot(emissions, "CO")
```

```{r}
quick_plot(emissions, "NH3")
```

```{r}
quick_plot(emissions, "NMVOC")
```

```{r}
quick_plot(emissions, "SO2")
```


It looks like there might not be any CO2 data from the CMIP6 era which seems off... but actually 
I am wondering if it would go into the LULC emissions and not the ffi emissions! 

```{r}
quick_plot(emissions, "CO2")
```



SO2 are in the same units... 

```{r}
quick_plot(emissions, "SO2")
```



N2O are also in the same units... 


```{r}
emissions %>% 
  filter(variable == "N2O") %>% 
  mutate(value = ifelse(source == "cmip7", value * 0.001, value))  %>% 
quick_plot(., "N2O")
```


What about NOx??? 

```{r}
quick_plot(emissions, "NOx")
```


```{r}
# Okay YES there is a difference in the NOx emissions... 
emissions %>% 
  filter(variable == "NOx") %>% 
  select(year, variable, value, source) %>% 
  spread(source, value) %>% 
  filter(year <= 2013) %>% 
  mutate(diff = cmip7/cmip6) %>% pull(diff) %>% mean()

# looks like the difference in the ratio is 1.535689.... 

```